---
title: "Homework 2"
author: "Zichen Shu"
output: github_document
---

```{r, message = FALSE}
library(tidyverse)
library(readxl)
library(dplyr)
```

## Problem 1

Read the Mr. Trashwheel dataset.

```{r, warning = FALSE}
trashwheel_df = 
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

```{r, message = FALSE, warning = FALSE}
precip_2018 =
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)
```

```{r}
precip_2017 =
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Now combine annual precipitation

```{r}
month_df =
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df = 
  bind_rows(precip_2018, precip_2017)

precip_df2 = 
  left_join(precip_df, month_df, by = "month")
```

This dataset contains indormation from the Mr. Trash Wheel collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data.

The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`

The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.



## Problem 2

```{r, warning = FALSE, message = FALSE}
subway_df = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name, route1:route11, station_latitude, station_longitude, entry, vending, entrance_type, ada) %>% 
  mutate(entry = recode(entry, "YES" = "TRUE", "NO" = "FALSE"), entry = as.logical(entry))

```

The dataset contains these **`r names(subway_df)`** variables. It consists of **`r nrow(subway_df)`** rows and **`r ncol(subway_df)`** columns. Some of the cleaning steps I have done so far are converting all variable names to lower case and replace the space to underscore, select the required variables and change the "entry" variable to a logical variable.

```{r}
###station_number = distinct(subway_df, station_number)



```











